### 2015-02-25 writing/F'14 paper/abstract.tex
Experimental setup tweaks; add TODOs to paper
Paper presentation overhaul.

\begin{abstract} 
  Recursive neural networks (RNNs) for sentence meaning have been
  successful for many tasks, but it remains an open question whether
  the fixed-length representations that they learn can support tasks as demanding as logical deduction.
  We pursue this question by evaluating whether two such models---plain RNNs and recursive neural
  tensor networks (RNTNs)---can correctly learn to identify logical
  relationships such as entailment and contradiction using these representations. 
  In our first set of experiments, we generate artificial data from a
  logical grammar and use it to evaluate the models' ability to learn
  to handle basic relational reasoning, recursive structures, and
  quantification. We then evaluate the models on the more natural
  SICK challenge data. Both models perform competitively on the SICK
  data and generalize well in all three experiments on simulated data,
  suggesting that they can learn suitable representations for logical
  inference in natural language. 
\end{abstract}

>>>>

\begin{abstract} 
  Tree-structured recursive neural networks (TreeRNNs) for sentence meaning have been
  successful for many tasks, but it remains an open question whether
  the fixed-length representations that they learn can support tasks as demanding as logical deduction.
  We pursue this question by evaluating whether two such models---plain TreeRNNs and recursive neural
  tensor networks (TreeRNTNs)---can correctly learn to identify logical
  relationships such as entailment and contradiction using these representations. 
  In our first set of experiments, we generate artificial data from a
  logical grammar and use it to evaluate the models' ability to learn
  to handle basic relational reasoning, recursive structures, and
  quantification. We then evaluate the models on the more natural
  SICK challenge data. Both models perform competitively on the SICK
  data and generalize well in all three experiments on simulated data,
  suggesting that they can learn suitable representations for logical
  inference in natural language.

  % TODO: Rewrite latter part, w/ or w/o SICK.
\end{abstract}

### 2015-02-25 writing/F'14 paper/discussion.tex
Experimental setup tweaks; add TODOs to paper
Paper presentation overhaul.

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then show that the same models can learn to
perform an entailment task on natural language image captions. The results suggest that RNTNs,
and potentially also RNNs, have the capacity to learn to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Of course, challenges remain. Even
the RNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or optimization procedures. The space of possible stronger
models is sizable, including extensions of the recursive models used here
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
models \cite{sutskever2014sequence}, and even learning-enabled variants 
of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
In addition, interesting analytical questions remain about \ii{how} these models encode
the underlying logics. Neither the underlying
logical theories, nor any straightforward parameter inspection technique provides 
much insight on this point, but we hope that further experiments may reveal 
some structure in the representations.

Our artificial data experiments have only scratched the surface 
of the logical complexity of natural language; in future experiments, we hope to test sentences
with more complex types of structure, as well as run similar experiments on non-tree structured models, such as LSTMs. Our SICK experiments 
similarly only begin to reveal the potential of these models to learn to 
perform complex semantic inferences from corpora. Nonetheless, the
rapid progress the field has made with these models in recent years
provides ample reason to be optimistic that they can be trained to
meet all the challenges of natural language semantics.

% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators. This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic. The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning. We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators. We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.

% There is ample room to build on these results. In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses. In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (-), or convolutional RNNs \cite{kalchbrenner2014convolutional}.

>>>>

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then show that the same models can learn to
perform an entailment task on natural language image captions. The results suggest that TreeRNTNs,
and potentially also TreeRNNs, have the capacity to learn to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Some questions about the abilities of these models remain open. Even
the TreeRNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or learning techniques. In addition, interesting 
analytical questions remain about \ii{how} these models encode
the underlying logics. Neither the underlying
logical theories, nor any straightforward parameter inspection technique provides 
much insight on this point, but we hope that further experiments may reveal 
some structure in the representations.

Our SICK experiments similarly only begin to reveal the potential of these models to learn to 
perform complex semantic inferences from corpora, and there is ample room to develop our understanding
using new sources of natural language data. Nonetheless, the rapid progress the field 
has made with these models in recent years provides ample reason to be optimistic that 
learned representation models can be trained to
meet all the challenges of natural language semantics.

### 2015-02-25 writing/F'14 paper/intro.tex
Paper presentation overhaul.

% TODO: TreeRNN.
% TODO: More language about the representations being the object of study.
% TODO: Is there any way to squeeze a useful example into the first page?
% TODO: Clarify early that we study vector pair classification. Maybe move up fig. 1?
% TODO: Clarify that we aren't implementing NatLog rules.
% TODO: ``Something that does that'' epigram in intro.
% TODO: Mention tree pair archichture.

Supervised recursive neural network models (RNNs; \citealt{goller1996learning}) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,irsoydeep},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn compositional semantic grammars, though many remain skeptical of the extent to which vector space representations can model complex compositional semantics, and it remains an open question whether learned can achieve the
high-fidelity distributed representations of recent algebraic
approaches to vector space modeling \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
and whether they can match the performance of grammars based in logical forms
in their ability to model core semantic phenomena like quantification, entailment, and contradiction \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}.

% TODO: Condense `skeptical' sentence

Recent work on the algebraic approach of
\newcite{ClarkCoeckeSadrzadeh2011} has yielded rich frameworks 
for computing the meanings of complex fragments of natural
language compositionally from vector or tensor representations, but 
has not yet yielded effective methods for learning these representations
entirely from data in a typical machine learning settings. 
To date, experimental investigations of these concepts using
distributed representations have been largely confined to short
phrases \cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
However, for robust natural language understanding, it is essential to
model these phenomena in their full generality on complex linguistic
structures. 
% TODO: Note about how their reps are learned instead.

Our research uses the task of \ii{natural language inference} 
(also known as \ii{recognizing textual entailment};
\citealt{dagan2006pascal}), in which the goal is to determine the core
inferential relationship between two sentences. Much of the
theoretical work on this task (and some successful implemented models;
\citealt{maccartney2009extended,watanabe2012latent}) involves \ii{natural
  logics}, which are formal systems that define rules of inference
between natural language words, phrases, and sentences without the
need of intermediate representations in an artificial logical
language. We use the powerful
 natural logic developed by
\newcite{maccartney2009extended} as our formal model. This logic
defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
and mutual consistency, as summarized in Table~\ref{b-table}, and it
provides rules of semantic combination for projecting these relations
from the lexicon up to complex phrases. The formal properties of this system are now well-understood
\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.

% TODO: Cut the below para. Replace with a more high-level summary of how we are paralleling NatLog, and how compositionality fits with that (Q: Does it?).

In our first set of experiments, we use this pre-specified logical
grammar to generate controlled data sets encoding the semantic
relationships between pairs of expressions and evaluate whether each
of two classes of neural model---plain recursive NNs and recursive neural
tensor networks (RNTNs; \citealt{socher2013acl1})---can learn those
relationships correctly. In our first experiment
(\S\ref{sec:join}), we evaluate the ability of these models to
learn the core relational algebra of natural logic from data. Our
second experiment (\S\ref{sec:recursion}) extends this
evaluation to relations between complex recursive structures like
$(a \plor b)$ and $\plneg (\plneg a \pland \plneg b)$, and our third
experiment (\S\ref{sec:quantifiers}) covers relations between
statements with quantification and negation like \ii{every reptile walks} and \ii{no turtle
 moves}. Though the performance of the plain RNN model 
 is somewhat poor in our first experiment, we find that the stronger RNTN model
  generalizes well in every case, suggesting that it has learned to simulate our target logical concepts. % TODO: Split into two sentences

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data. The small size of this corpus puts
data-hungry NN models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
surpassing several submitted models with significant hand-engineered task-specific features.
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that RNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

>>>>

% TODO: Is there any way to squeeze a useful example into the first page?
% TODO: Can we fit a version of the ``find something that does that'' quote in intro somewhere?

% TODO: This para might be a bit repetitive.
Tree-structured recursive neural network models (TreeRNNs; \citealt{goller1996learning}) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,irsoydeep},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn to produce and use strong 
semantic representations for sentences,
though many remain skeptical of the extent to which they can learn to represent
the necessary complex compositional structures using fixed-size vectors. 
In particular, it remains an open question whether learned models can achieve the
high-fidelity distributed representations of recent algebraic
approaches to vector space modeling \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
and whether they can match the performance of grammars based in logical forms
in their ability to model core semantic phenomena like quantification, entailment, and contradiction \cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}.

Recent work on the algebraic approach of
\newcite{ClarkCoeckeSadrzadeh2011} has yielded rich frameworks 
for computing the meanings of complex fragments of natural
language compositionally from vector or tensor representations, but 
has not yet yielded effective methods for learning these representations
entirely from data in typical machine learning settings. 
To date, experimental investigations of these concepts using
distributed representations have been largely confined to short
phrases \cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
However, for robust natural language understanding, it is essential to
model these phenomena in their full generality on complex linguistic
structures.

This paper describes four machine learning experiments
that directly evaluate the abilities of these models to learn representations that 
support specific semantic behaviors. These tasks follow the format of
 \ii{natural language inference} 
(also known as \ii{recognizing textual entailment};
\citealt{dagan2006pascal}), in which the goal is to determine the core
inferential relationship between two sentences. 
We introduce a novel NN architecture for natural language inference which
independently computes vector representations for each of two sentences using
standard TreeRNN or TreeRNTN \cite{socher2013acl1} models, and
produces a judgment for the pair using only those representations. This allows
us to gauge the abilities of these two models to represent all of the necessary semantic
information in the sentence vectors.

Much of the
theoretical work on natural language inference (and some successful implemented models;
\citealt{maccartney2009extended,watanabe2012latent}) involves \ii{natural
  logics}, which are formal systems that define rules of inference
between natural language words, phrases, and sentences without the
need of intermediate representations in an artificial logical
language. This logic
defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
and mutual consistency, as summarized in Table~\ref{b-table}, and it
provides rules of semantic combination for projecting these relations
from the lexicon up to complex phrases. The formal properties of this system 
are now well-understood \cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.

While one of our experiments uses natural logic only as an approximate
description of the desired model behavior, three experiments draw on this
 tradition much more directly, and
test models' abilities to learn to reproduce the precise behavior of the natural logic developed by
\newcite{maccartney2009extended}: one
covers learning to reason with the bare logical relations (\S\ref{sec:join}), a second covers reasoning with statements constructed compositionally from recursive functions (\S\ref{sec:recursion}),
and a third covers the additional complexity that results from quantification (\S\ref{sec:quantifiers}).
Though the performance of the plain TreeRNN model 
 is somewhat poor in our first experiment, we find that the stronger TreeRNTN model
  generalizes well in every case, suggesting that it has learned to simulate our target logical concepts.

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data in \S\ref{sec:sick}. The small size of this corpus puts
data-hungry NN models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
surpassing several submitted models with significant hand-engineered task-specific features.
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that TreeRNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

### 2015-02-25 writing/F'14 paper/join.tex
Paper presentation overhaul.

@@

 % TODO: Why is it okay to study non-recursive models here.
 % TODO: Run a baseline for shits and giggles?
 % TODO: Mention that we do an easier version of this implicitly elsewhere.

If a model is to learn the behavior of a relational logic like ours
from a finite amount of data, it must learn to deduce new
relations from seen relations in a sound manner. The simplest such
deductions involve atomic statements using the relations in
Table~\ref{b-table}. 
For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
one can conclude that $p_2 \natfor p_1$ using the definitions of the relations directly. 
If one is also given the relation $p_1 \natrev p_3$, one can conclude that $p_1 \natrev p_2$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$p_1 \natfor p_2$ and $p_2 \natneg p_3$, it follows that $p_1 \natalt p_3$.  The
full set of sound such inferences on pairs of premise relations is depicted in
Table~\ref{tab:jointable}.

@@

\begin{table}[htp]
  \centering  \small
  \setlength{\arraycolsep}{8pt}
  \renewcommand{\arraystretch}{1.1}
  \newcommand{\UNK}{\cdot}  
  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
    %\hline
    \multicolumn{1}{c}{}
             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
    \cline{2-8}
    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
    \cline{2-8}
  \end{array}$
  \caption{Inference path from premises $p_1Rp_2$ (row) and $p_2Sp_3$ (column) to the relation that holds between $p_1$ and $p_3$, if any.  These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}. We assess our models' ability to reproduce such inferential paths. Cells containing a dot correspond to pairs
of relations for which no valid inference can be drawn.} % TODO: Move harder example in to replace caption.
  \label{tab:jointable}
\end{table}

\paragraph{Experiments}
To generate data for these inferential patterns, we
create small boolean structures for our logic in which terms denote
sets of entities from a small domain.  Figure~\ref{lattice-figure}
depicts a structure of this form. The lattice gives the full model,
and we generate a relational statement for each pair of labels in the model. 
We divide these statements evenly into training and test sets, and remove from the
test set statements which cannot be proven from the training
statements.
In each experimental run, we create 80 randomly generated sets drawing from
a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
2960 relations.
% TODO: Explain what the lattice shows.

We trained models with both the NN and NTN comparison functions on these
data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no strictly simpler model is likely viable as a baseline.} In both cases, the models are implemented as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not used, and the two models are not recursive. We simply present
the models with the (randomly initialized) embedding vectors for each
of two terms, ensuring that the model has no information about the terms
being compared except for the relations between them that appear in training.

@@

      \labelednode{2.5}{0.5}{}{}
    \end{picture}}
    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
  \end{subfigure}
  \qquad
    \begin{subfigure}[t]{0.43\textwidth}
    \centering \vspace{0.4cm}
    \setlength{\tabcolsep}{12pt}
    \begin{tabular}[b]{c  c}
      \toprule
      Train & Test \\
      \midrule
      $p_1 \nateq p_2$              & $p_2 \natneg p_7$ \\
      $p_1 \natrev p_5$              & $p_2 \natrev p_5$ \\
      $p_4 \natrev p_8$              & \strikeout{$p_5 \nateq p_6$} \\
      $p_5 \natalt p_7$              & \strikeout{$p_7 \natfor p_4$} \\
      $p_7 \natneg p_1$           & $p_8 \natfor p_4$ \\

@@

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{15d RNN}  & \multicolumn{2}{c}{15d RNTN}\\
    \midrule
    Train &53.8 &(10.5)    & 99.8&	(99.0) & \textbf{100} & \textbf{(100)}\\
    Test &52.0 &(10.3) &	94.0&(87.0)& \textbf{99.6} & \textbf{(95.5)}\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the semantic relation experiments. These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
  \label{joinresultstable}
\end{table}

@@

In one run for example, both models successfully classified the test example $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$. On a somewhat similar example involving comparably frequent relations, the RNTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the RNN incorrectly assigned it $\natind$.

>>>>

\begin{table}[tp]
  \centering  \small
  \setlength{\arraycolsep}{8pt}
  \renewcommand{\arraystretch}{1.1}
  \newcommand{\UNK}{\cdot}  
  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
    %\hline
    \multicolumn{1}{c}{}
             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
    \cline{2-8}
    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
    \cline{2-8}
  \end{array}$
  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of relations using the rules represented here. As an example, given that $p_1 \natfor p_2$ and $p_2 \natneg p_3$, the entry in the $\natfor$ row and the $\natneg$ column lets us conclude that $p_1 \natalt p_3$.
These rules are derived from the definitions of the relations in Table~\ref{b-table}.  Cells containing a dot correspond to pairs
of relations for which no valid inference can be drawn.} 
  \label{tab:jointable}
\end{table}

@@

The simplest kinds of deduction in natural logic involve atomic statements 
using the relations in Table~\ref{b-table}. 
For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
one can conclude that $p_2 \natfor p_1$ using the definitions of the relations directly. 
If one is also given the relation $p_1 \natrev p_3$, one can conclude that $p_1 \natrev p_2$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
full set of sound such inferences on pairs of premise relations is depicted in
Table~\ref{tab:jointable}. Though these basic inferences do not involve compositional
sentence representations, any successful reasoning using compositional representations
will rely on the ability to perform sound inferences of this kind, so our first experiment studies how well each model can learn to perform them them in isolation.

@@

\paragraph{Experiments}
We begin by creating a world model
on which we will base the statements in the train and test sets.
This takes the form of a small boolean structure in which terms denote
sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight terms ($p_{1\text{-}8}$). We then generate a 
relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
We divide these statements evenly into training and test sets, and remove from the
test set statements which cannot be proven from the training
statements.
In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
2960 relations.

We trained models with both the NN and NTN comparison functions on these
data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} In both cases, the models are implemented as
described in \S\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not used, and the two models are not recursive. We simply present
the models with the (randomly initialized) embedding vectors for each
of two terms, ensuring that the model has no information about the terms
being compared except for the relations between them that appear in training.

@@

      \labelednode{2.5}{0.5}{}{}
    \end{picture}}
    \caption{Example boolean structure. The terms $p_{1\text{-}8}$ name the sets. Not all sets have names, and
    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
  \end{subfigure}
  \qquad\small
    \begin{subfigure}[t]{0.43\textwidth}
    \centering \vspace{0.4cm}
    \setlength{\tabcolsep}{12pt}
    \begin{tabular}[b]{c  c}
      \toprule
      Train & Test \\
      \midrule
      $p_1 \nateq p_2$              & $p_2 \natneg p_7$ \\
      $p_1 \natrev p_5$              & $p_2 \natrev p_5$ \\
      $p_4 \natrev p_8$              & \strikeout{$p_5 \nateq p_6$} \\
      $p_5 \natalt p_7$              & \strikeout{$p_7 \natfor p_4$} \\
      $p_7 \natneg p_1$           & $p_8 \natfor p_4$ \\

@@

\begin{table}[tp]
  \centering\small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{Train} & \multicolumn{2}{c}{Test}\\
    \midrule
    $\natind$ only &53.8 &(10.5)    &53.8 &(10.5) \\
    15d NN &				99.8&	(99.0) &94.0&(87.0) \\
    15d NTN 				& \textbf{100} & \textbf{(100)} & \textbf{99.6} & \textbf{(95.5)}\\
    \bottomrule
  \end{tabular}

  \caption{Performance on the semantic relation experiments. These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
  \label{joinresultstable}
\end{table}

@@

As an example, both models successfully classified the test item $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$. On another example involving comparably frequent relations, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.

### 2015-02-25 writing/F'14 paper/methods.tex
Paper presentation overhaul.

We limit the scope of our experiments in this paper neural network models 
that adhere to the linguistic \ii{principle of
 compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their parts
via specific composition functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are vectors of dimension $n$. The
composition function maps pairs of them to single vectors of dimension $n$, 
which can then be merged again to represent more complex
phrases. Once the entire sentence-level representation has been
derived, it serves as a fixed-dimensional input for some subsequent function.

% TODO: Say that there are two RNTNs in caption.

\begin{figure}[tp]
  \centering
 % \includegraphics[scale=0.35]{model.eps}
  \input{figure1}
  \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
    The same structure is used for both the RNN and RNTN layer functions.} 
  \label{sample-figure}
\end{figure}

To apply these recursive models to our task, we propose the tree 
pair model architecture depicted in Figure~\ref{sample-figure}. 
In it, the two phrases being compared are processed separately using a pair 
of recursive networks that share a single set of parameters. 
The resulting vectors are fed into a separate comparison
layer that is meant to generate a feature vector (fixed at 75 dimensions) capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which in turn produces a hypothesized
distribution over the seven relations represented in Table~\ref{b-table}.
% TODO: Move 75 constant later.

For a composition layer, we evaluate models with both the plain neural
network layer function \eqref{rnn} and the more powerful RNTN layer function
\eqref{rntn} proposed in \newcite{chen2013learning}. The nonlinearity $f(x) = \tanh(x)$ is applied to the output of the layer function.
%
\begin{gather} 
\label{rnn}
\vec{y}_{\textit{TreeRNN}} = f(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{rntn} 
\vec{y}_{\textit{TreeRNTN}} = \vec{y}_{\textit{TreeRNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather} 
%
Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
representations for the left and right children of the node, and
$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
them by an $n \times 2n$ matrix of learned weights, adds a bias $\vec{b}$, and applies the
non-linearity element-wise to the resulting vector. The RNTN adds a learned third-order tensor 
$\mathbf{T}$, of dimension $n \times n \times n$, modeling
multiplicative interactions between the child vectors.

The comparison layer uses the same layer function as the
composition layers (either an NN layer or an NTN layer) with
independently learned parameters and a separate nonlinearity function.
Rather than use a $\tanh$ nonlinearity here, we found better results with a rectified linear function. In
particular, we use the leaky rectified linear function
\cite{maasrectifier}: $f(x)=\max(x, 0) +
0.01\min(x, 0)$.

To run the model forward and label a pair of phrases, the structure of
the lower layers of the network is assembled so as to match the tree
structures provided for each phrase, which is either included in the source data or given by a parser.
The word vectors are then looked up from the vocabulary matrix $V$ (one of the model parameters), and
the composition and comparison functions are used to pass information
up the tree and into the classifier at the top. For an objective
function, we use the negative log of the probability assigned to the
correct label with L2 regularization.

% Removed 'minibatch' -> For two of the experiments, we use minibatches
% of size 1, which doesn't really count.
We train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
The hyperparameters of AdaDelta, the dimensions of the representations,
and the initialization strategy are all tuned manually. Training times on CPUs vary from hours to days across experiments.
Because the classes are not balanced in the artificial data experiments, we report performance
using both accuracy and macroaveraged F1.\footnote{We compute macroaveraged F1 
as the harmonic mean of average precision and average recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined.} On the artificial data experiments, we report mean
results over five fold cross-validation. Variance across runs is typically no
more than two percentage points.
Source code and generated data will be released after the review period.

>>>>

We limit the scope of our experiments in this paper to neural network models 
that adhere to the linguistic \ii{principle of
 compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their parts
via specific composition functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are embedding vectors of dimension $n$. The
composition function maps pairs of them to single phrase vectors of dimension $n$, 
which can then be merged again to represent more complex
phrases. Once the entire sentence-level representation has been
derived, it serves as a fixed-dimensional input for some subsequent function.

\begin{figure}[tp]
  \centering
  \input{figure1}
  \caption{In our model, two separate tree-structured networks build up vector representations for each of two sentences using either NN or NTN layer functions. A comparison layer then uses the resulting vectors to produce features for a classifier.} 
  \label{sample-figure}
\end{figure}

To apply these recursive models to our task, we propose the tree 
pair model architecture depicted in Fig.~\ref{sample-figure}. 
In it, the two phrases being compared are processed separately using a pair 
of tree-structured networks that share a single set of parameters. 
The resulting vectors are fed into a separate comparison
layer that is meant to generate a feature vector capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which produces a
distribution over the seven relations represented in Table~\ref{b-table}.

For the sentence embedding portions of the network, we evaluate both TreeRNN models with
 the standard NN layer function \eqref{TreeRNN} and those with the more powerful neural tensor 
 network layer function
\eqref{TreeRNTN} proposed in \newcite{chen2013learning}. The nonlinearity $f(x) = \tanh(x)$ is applied
 elementwise to the output of either layer function.
%
\begin{gather} 
\label{TreeRNN}
\vec{y}_{\textit{TreeRNN}} = f(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{TreeRNTN} 
\vec{y}_{\textit{TreeRNTN}} = \vec{y}_{\textit{TreeRNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather} 
%
Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
representations for the left and right children of the node, and
$\vec{y}$ is the node's output.  The TreeRNN concatenates them, multiplies
them by an $n \times 2n$ matrix of learned weights, and adds a bias $\vec{b}$. 
The TreeRNTN adds a learned third-order tensor 
$\mathbf{T}$, of dimension $n \times n \times n$, modeling
multiplicative interactions between the child vectors. 
Other tree-structured networks have been successfully used in past work
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
we believe that these two provide a valuable case study, and that positive results on 
here are likely to generalize well to stronger models.

The comparison layer uses the same layer function as the
composition layers (either an NN layer or an NTN layer) with
independently learned parameters and a separate nonlinearity function.
Rather than use a $\tanh$ nonlinearity here, we found better results with the leaky rectified linear function
\cite{maasrectifier}: $f(x)=\max(x, 0) +
0.01\min(x, 0)$.

To run the model forward and label a pair of phrases, the structure of
the lower layers of the network is assembled so as to match the tree
structures provided for each phrase, which is either included in the source data or given by a parser.
The word vectors are then looked up from the  vocabulary embedding matrix $V$ (one of the model parameters), and
the composition and comparison functions are used to pass information
up the tree and into the classifier at the top. For an objective
function, we use the negative log of the probability assigned to the
correct label with L2 regularization.

We initialize parameters uniformly, using the range $(-0.05, 0.05)$ for layer parameters
and $(-0.01, 0.01)$ for embeddings, and train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
The classifier feature vector is fixed at 75 dimensions and 
the dimensions of the recursive layers are tuned manually.
Training times on CPUs vary from hours to days across experiments.
On the first three experiments, which use artificial data, we report mean
results over five fold cross-validation.
Variance across runs is typically no
more than two percentage points. In addition, because the classes are not necessarily balanced, 
we report performance 
using both accuracy and macroaveraged F1.\footnote{We compute macroaveraged F1 
as the harmonic mean of average precision and average recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined.}
Source code and generated data will be released after the review period.
